\newgeometry{left=5cm}

\section*{Acknowledgements}
\label{sec:acknowledgements}

This work would not be possible to be accomplished without effort, help and support from my family, fellows and colleagues. Thus, in this section I would like to give my sincere thanks to all of them.

Starting by giving thanks to my mother and to my whole family, who have supported me through this entire and long journey, and who always gave and will always give me some of the most important and beautiful things in life, love and friendship.

In second place, I would like to thank all people that were involved directly in this project. To my supervisor, Professor Filipe Araújo, who contributed with his vast wisdom and experience, to my co-supervisor, Professor Jorge Cardoso, who contributed with is vision and guidance about the main road we should take and to Engineer Jaime Correia, who ``breathes'' these kind of topics through him and helped a lot with is enormous knowledge and enthusiasm.

In third place, I would like to thank Department of Informatics Engineering and the Centre for Informatics and Systems, both from the University of Coimbra, for allowing and provide the resources and facilities to  to be carried out this project.

In fourth place, to the Foundation for Science and Technology (FCT), for financing this project facilitating its accomplishment, to Huawei, for providing tracing data, core for this whole research, and to Portugal National Distributed Computing Infrastructure (INCD) for providing hardware to run experiments.

And finally, my sincere thanks to everyone that I have not mentioned and contributed to everything that I am today.

\restoregeometry